---
title: "Data Quality API Import Specification and Payload Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data Quality API Import Specification and Payload Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette summarizes `data_quality_api` import (`page=import`) behavior and maps it
to recommended inputs for `redcapdqapi::dq_import()`. It focuses on:

1. Which fields are needed to create a new query row.
2. Which fields are needed to add a comment to an existing query.

## Import endpoint and core constraints

`dq_import()` calls the REDCap API with module parameters:

- `type=module`
- `prefix=vanderbilt_dataQuality` (override with `dq_client(..., prefix=)`)
- `page=import`

Important import behavior in `data_quality_api`:

- Import only adds new comments/replies; it does not modify existing comments.
- Deduplication is based on `ts + user` (resolved to `user_id`) within a query; duplicates are skipped.
- `data` must be provided as JSON text.

## Top-level payload structure

`data` must represent a collection of query rows. Two JSON shapes are accepted:

- object: `{"123": {...}, "124": {...}}`
- array: `[{...}, {...}]`

`dq_import()` accepts three input forms:

- raw JSON string
- export-shaped R list (automatically converted using `toJSON`)
- minimal data frame (converted by the package into OPEN-query payload)

## Fields required to create a new query row

When no existing query is matched, the API attempts to create a new query row. To succeed,
`dataRow` should satisfy:

- `project_id` must match the project associated with the API token.
- `record` must be non-empty.
- `event_id` must be valid (single-event projects may auto-apply event in some cases).
- `field_name` must exist in project metadata.

Recommended fields in practice:

- required: `project_id`, `record`, `event_id`, `field_name`
- strongly recommended: `status_id` (can be empty string), `instance`, `repeat_instrument`, `assigned_username`

> Note: `repeat_instrument` may participate in status matching, but may not be persisted in the
> query master row.

## Fields required to add a comment to an existing query

Each comment is in `dataRow$resolutions[[i]]`. Recommended minimum fields:

- `ts`: timestamp string
- `username`: REDCap username resolvable to a valid user
- `comment`: comment text

Query-state fields (recommended):

- `current_query_status`: e.g., `OPEN`, `CLOSED`, `VERIFIED`, `DEVERIFIED`
- `response_requested`: `0` or `1`
- `response`: e.g., `DATA_MISSING`, `TYPOGRAPHICAL_ERROR`, `CONFIRMED_CORRECT`, `WRONG_SOURCE`, `OTHER`

For existing queries, the most stable approach is to set the same existing status ID in both
`dataRow$status_id` and `resolution$status_id`.

## Minimal payload examples

### A. Create a new query + one comment

```{r}
new_status_payload <- list(
  list(
    status_id = "",
    project_id = "123",
    record = "A001",
    event_id = "1",
    field_name = "age",
    instance = NULL,
    repeat_instrument = NULL,
    assigned_username = NULL,
    resolutions = list(
      list(
        res_id = "",
        status_id = "",
        ts = "2026-02-18 13:45:00",
        username = "jubilee",
        comment = "Please verify age value.",
        response_requested = 1,
        response = NULL,
        current_query_status = "OPEN"
      )
    )
  )
)

jsonlite::toJSON(new_status_payload, auto_unbox = TRUE, null = "null", pretty = TRUE)
```

### B. Add one comment to an existing query

```{r}
existing_status_payload <- list(
  list(
    status_id = "9999",
    project_id = "123",
    record = "A001",
    event_id = "1",
    field_name = "age",
    instance = NULL,
    repeat_instrument = NULL,
    assigned_username = NULL,
    resolutions = list(
      list(
        res_id = "",
        status_id = "9999",
        ts = "2026-02-18 14:00:00",
        username = "jubilee",
        comment = "Follow-up: confirm source document.",
        response_requested = 0,
        response = "OTHER",
        current_query_status = "OPEN"
      )
    )
  )
)

jsonlite::toJSON(existing_status_payload, auto_unbox = TRUE, null = "null", pretty = TRUE)
```

## Calling `dq_import()`

```{r eval = FALSE}
library(redcapdqapi)

client <- dq_client(
  api_url = "https://redcap.example.org/api/",
  token = Sys.getenv("REDCAP_TOKEN"),
  pid = 123
)

# Import directly from list input (automatically converted to JSON)
dq_import(client, existing_status_payload)
```

## Security notes

- Store API tokens in environment variables (for example, `REDCAP_TOKEN`).
- Never commit tokens to version control.
- If import fails, first verify required field completeness and `project_id / status_id` alignment for the target query.

## Limitations and upstream constraints

### Repeat instrument handling

The REDCap Data Quality API `import.php` implementation does not store the `repeat_instrument`
column when inserting new rows into `redcap_data_quality_status`.

Although `repeat_instrument` is used when attempting to match existing query records,
it is not written during status creation. As a result:

- A new query cannot be created that is scoped specifically to a repeating instrument.
- Queries created via the API will have `repeat_instrument = NULL`.
- In repeating instrument contexts, instance-level isolation may not behave as expected.

This behavior originates from the upstream REDCap Data Quality External Module
and is not controlled by `redcapdqapi`.

Users who require repeat-instrument–aware query creation must modify the upstream
`import.php` implementation accordingly.
